---
title: Simulate from bootstrap estimates
author: ""
output: 
  github_document:
    toc: true
---


```{r, echo = FALSE}
knitr::opts_chunk$set(comment = '.', fig.path = "img/bootstrap-")
```

# Scope and source code


This document demonstrates how to use an mrgsolve model and bootstrap 
parameter estimates to create replicate simulations that incorporate
uncertainty in the fixed effect (e.g. `THETA`) and random effect (e.g.
`OMEGA` and `SIGMA`) parameters.

__Source code__ 

The example code for this document can be reached [here](bootstrap.Rmd) and 
the example model [here](bootstrap.cpp).  The bootstrap data set is bundled
with the mrgsolve package.


__Required packages__
```{r, message = FALSE}
library(mrgsolve)
library(dplyr)
library(ggplot2)
```


# Load an example model

You can see the model source [here](bootstrap.cpp)
```{r, message = FALSE}
mod <- mread("bootstrap")
```

We have `THETA1`, `THETA2`, and `THETA3` in `$PARAM`

```{r}
param(mod)
```

We have a 3x3 `$OMEGA` matrix
```{r}
omat(mod)
```

and a 1x1 `$SIGMA` matrix
```{r}
smat(mod)
```

# Generate an example data template

```{r}
data <- ev_rep(ev(amt = 100, ii = 24, addl = 2), ID = 1:5)
```

100 mg daily x3 in 5 individuals

```{r}
data
```


## Test simulation


The basic simulation we will be doing is
```{r}
set.seed(987652)

mod %>%
  data_set(data) %>%
  mrgsim(Req = "DV", end = 96)  %>% 
  plot()
```

Just three doses, simulating out to 96 hours.  From here, we will do this 
test simulation for several replicates, with each replicate drawing a different
set of bootstrap parameters.

# Example bootstrap output

The example assumes that the different estimates are in 
columns and the different boostrap runs are each 
in a separate row of a data frame.

```{r}
data(exBoot)
head(exBoot)
```


__NOTE__ mrgsolve has some
 functions to help you deal with 
output that is in this `NONMEM` type format; for 
`OMEGA`, it is expecting `OMEGA11`, then `OMEGA21`
then `OMEGA22`  then `OMEGA31` etc.  Below, find
some functions to go into this data frame to 
make R matrices that you can pass to the 
model. 

I know that PsN can return these data in a 
different order.  I don't know enough about 
how it is structured to write something that 
will always handle properly and in a convenient
way.  So there if you're using PsN, there is a chance
you will have to code some of this by hand.  If that
is a case, I'm looking for a PsN-using collaborator
who could help us understand / predict the 
output so we could automate. 


# Helper functions for matrices

## Create a list of `$OMEGA` and `$SIGMA` matrices

Create a list of `OMEGA` matrices; the function looks for column
names in `exBoot` that contains `OMEGA`
```{r}
omegas <- as_bmat(exBoot, "OMEGA")
```

```{r}
omegas[[10]]
```

```{r}
length(omegas)
```

You can also just go after a single row
```{r}
as_bmat(exBoot[1,], "OMEGA")
```


Do this for `SIGMA` too; I only had one `EPS` in the 
simulation model, so I'm going to look for `SIGMA11` only
to get a 1x1 matrix
```{r}
sigmas <- as_bmat(exBoot, "SIGMA11")

sigmas[[10]]
```

# Simulate

1.  The `param` call scrapes the `THETA`s from `exBoot`; this works
because `THETA1`, `THETA2`, and `THETA3` are listed in `$PARAM`
1.  Use `omat` to update the `$OMEGA` matrix; this works because
`$OMEGA` is a 3x3 matrix
1.  Use `smat` to update the `$SIGMA` matrix

In the simulation loop, `i` indicates the replicate number, or the `ith`
set of bootstrap parameter estimates.

```{r}
set.seed(222)

out <- lapply(1:10, function(i) {
  
  mod %>%
    param(exBoot[i,]) %>%
    omat(omegas[[i]]) %>%
    smat(sigmas[[i]]) %>%
    data_set(data) %>%
    mrgsim(Req = "DV", end = 96) %>%
    mutate(rep = i)
  
}) %>% bind_rows
```


In the output, we have 10 replicates, each with five individuals
```{r}
ggplot(out, aes(time,DV,group=ID)) + geom_line() + theme_bw() + facet_wrap(~rep)
```

# Simulate with only uncertainty in the `THETA`s

Here, we just drop out the update to `OMEGA` and `SIGMA` and 
zero out the random effects.

```{r}
set.seed(222)

dat1 <- filter(data, ID==1)

out <- lapply(1:10, function(i) {
  
  mod %>%
    param(exBoot[i,]) %>%
    zero_re() %>%
    data_set(dat1) %>%
    mrgsim(Req = "DV", end = 96) %>%
    mutate(rep = i)
  
}) %>% bind_rows
```

Now, we have one "individual" simulated from 10 different bootstrap 
parameter sets

```{r}
ggplot(out, aes(time,DV,group=rep)) + geom_line() + theme_bw() 
```

# Session Info
```{r}
devtools::session_info()
```

