---
title: Simulate from bootstrap estimates
author: ""
output: 
  github_document:
    toc: true
---


```{r, echo = FALSE}
knitr::opts_chunk$set(comment = '.', fig.path = "img/bootstrap-")
```



```{r, message = FALSE}
library(mrgsolve)
library(dplyr)
library(ggplot2)
```

# Load an example model

You can see the model source [here](bootstrap.cpp)
```{r, message = FALSE}
mod <- mread("bootstrap")
```

# Example bootstrap output

The example assumes that the different estimates are in 
columns and the different boostrap runs are each 
in a separate row of a data frame.

```{r}
data(exBoot)
head(exBoot)
```


__NOTE__ mrgsolve has some
 functions to help you deal with 
output that is in this `NONMEM` type format; for 
`OMEGA`, it is expecting `OMEGA11`, then `OMEGA21`
then `OMEGA22`  then `OMEGA31` etc.  Below, find
some functions to go into this data frame to 
make R matrices that you can pass to the 
model. 

I know that PsN can return these data in a 
different order.  I don't know enough about 
how it is structured to write something that 
will always handle properly and in a convenient
way.  So there if you're using PsN, there is a chance
you will have to code some of this by hand.  If that
is a case, I'm looking for a PsN-using collaborator
who could help us understand / predict the 
output so we could automate. 


# Helper functions for matrices

## Create single `$OMEGA` matrix

This creates a matrix from data in the 
first row where the name contains `OMEGA`
```{r}
as_bmat(exBoot[1,], "OMEGA")
```

If you have to code something by hand, these might help
```{r}
bmat(c(1,2,3))

dmat(c(1,2,3))
```


## Create a list of `$OMEGA` and `$SIGMA` matrices

Create a list of `OMEGA` matrices
```{r}
omegas <- as_bmat(exBoot, "OMEGA")
```

```{r}
omegas[[10]]
```

```{r}
length(omegas)
```


Do this for `SIGMA` too; I only had one `EPS` in the 
simulation model, so I'm going to look for `SIGMA11` only
to get a 1x1 matrix
```{r}
sigmas <- as_bmat(exBoot, "SIGMA11")

sigmas[[10]]
```

# Simulate

## Update the parameters

The `param` call scrapes the `THETA`s from `exBoot`; this works
because `THETA1`, `THETA2`, and `THETA3` are listed in `$PARAM`
```{r}
param(mod)
```

## Upate the matrices

Use `omat` to update the `$OMEGA` matrix; this works because
`$OMEGA` is a 3x3 matrix
```{r}
omat(mod)
```



Use `smat` to update the `$SIGMA` matrix

```{r}
smat(mod)
```


## Simulation loop

```{r}
set.seed(222)

out <- lapply(1:10, function(i) {
  
  mod %>%
    param(exBoot[i,]) %>%
    omat(omegas[[i]]) %>%
    smat(sigmas[[i]]) %>%
    ev(amt = 100, ii = 24, addl = 2) %>%
    mrgsim(Req = "DV", end = 96) %>%
    mutate(rep = i)
  
}) %>% bind_rows
```


```{r}
ggplot(out, aes(time,DV,group=rep)) + geom_line()
```

# Simulate with only uncertainty in the `THETA`s

All the uncertainty comes from the bootstrap `THETA`s

```{r}
set.seed(222)

out <- lapply(1:10, function(i) {
  
  mod %>%
    param(exBoot[i,]) %>%
    zero_re() %>%
    ev(amt = 100, ii = 24, addl = 2) %>%
    mrgsim(Req = "DV", end = 96) %>%
    mutate(rep = i)
  
}) %>% bind_rows
```


```{r}
ggplot(out, aes(time,DV,group=rep)) + geom_line()
```

# Session Info
```{r}
devtools::session_info()
```

