---
title: "Interoccasion variability"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = '.', message = FALSE, fig.path = "img/iov-")
```

```{r}
library(mrgsolve)
library(dplyr)
```


## Model code

- More than likely, you will need a occasion indicator (like `OCC`) 
in both your parameter list and the data set.
- Create a variable (`IOV`) that holds the inter-occasion random effect.
- For three occasions, you'll need three ETAs; assign them as appropriate.
- Recall that `mrgsolve` uses `nocb` advance.
- For this set up, you'll need 4x4 omega block (one for CL, and three
for the occasions).
- There is no need for `SAME` as in NONMEM; you always have control over
the value for the OMEGAs.  Just make them the __same__ yourself. 
- You don't have to use __labels__ but it might be a good idea to 
avoid confusion.

```{r}
code <- '
$PARAM TVCL = 1, V = 20, KA = 1, OCC = 1

$PKMODEL cmt = "GUT CENT", depot = TRUE

$MAIN
capture IOV = ETA_OCC1;

if(OCC ==2) IOV = ETA_OCC2;

if(OCC ==3) IOV = ETA_OCC3;

capture CL  = TVCL*exp(ETACL + IOV);

$OMEGA @annotated
ETACL   : 0.1 : .
ETA_OCC1: 1.0 : .
ETA_OCC2: 1.0 : . 
ETA_OCC3: 1.0 : .

$POST
capture CP = CENT/V;

$SET req = ""
'
```

```{r}
mod <- mcode_cache("iov", code)
```

## The data set

```{r}
data <- realize_addl(ev(amt = 100, ii = 72, addl = 3, ID = 1)) %>% as.data.frame

data <- mutate(data, OCC = cumsum(evid)-1)
data
```


## And simulate ...
```{r}
set.seed(101)
mrgsim(mod, data=data, end = 480) %>% plot
```

