---
title: Mixture model example
output: github_document
---


```{r,echo=FALSE}
knitr::opts_chunk$set(message=FALSE, comment='.', fig.path = "img/mixture")
```


In NONMEM, `$MIXTURE` will allow you to estimate mixture models, where
individuals are classified in to two or more populations with 
a certain probability. It is straightforward to simulate from models like these in your `mrgsolve` model code.


# Two Populations
Let's imagine there were two populations in the mixture model, with 
the second having smaller clearance than the first.  In this example, 
we will develop some code for a simple model and then extend it
to implement the mixture model component.

A simple model might be:

```{r,message=FALSE}
library(mrgsolve)
library(dplyr)

code <- '
$PARAM TVCL = 1, V = 30, KA=1.2, POP = 1, THETA1 = 0.5

$PKMODEL cmt="GUT CENT", depot=TRUE

$MAIN
double CL = TVCL;

if(POP==2) CL = TVCL * THETA1;
'
```

In this model, we created a parameter for the population indicator
(`POP`) and if `POP` is 2 then clearance is lower than it would be 
otherwise.


Compile this model and run
```{r}
mod <- mcode_cache("A", code) %>% update(end=72, delta=0.1)

idata <- data_frame(POP=c(1,2))

e <- ev(amt=100)

mod %>% mrgsim(idata=idata,events=e) %>% plot
```

The profile in pink was for `POP==2` or the lower clearance profile and
blue was for `POP==1`.


## Modify the model to simulate a population mixture
In the get-started exmaple model, we hard-coded `POP` as a parameter and 
we had to supply the value of `POP` in the input data set (in this case,
it was via `idata`).

For the mixture model, we want `POP` to be simulated and we want the 
simulated value to be 1 with a probability of, say, 0.8 and 2 
with a probability of 0.2. 

To make this happen, we need to simulate a binary variate for 
each individual.  Random numbers are easy to simulate with `mrgsolve`
when you use `$PLUGIN`.

```{r}
code <- '
$PLUGIN Rcpp

$MAIN
if(NEWIND <=1) {
  int POP = 1 + R::rbinom(1,0.2);
}

$CAPTURE POP
'

mod <- mcode_cache("B", code)
```

Here, we invoked the `Rcpp` plugin that allows us to call `R::binom(1,0.8)`.  `R::binom` is just like the regular `R` version, but it only draws
one variate (`n=1`).

Let's test it out

```{r}
set.seed(222)
out <- mrgsim(mod, nid=10000, end=-1)

head(out)
```

Here, we've got 20% of the people in the population with `POP` of 2:
```{r}
mean(out$POP==2)
```

Now, let's modify the model again to incorporate our random `POP`
calculation with the PK model.  I have also included 
a home-brewed `ETA` using `R::rnorm` as another example and to 
make the summary a little more interesting.

```{r}
code <- '
$PLUGIN Rcpp

$PARAM TVCL = 1, V = 30, KA=1.2, THETA1 = 0.5

$PKMODEL cmt="GUT CENT", depot=TRUE

$MAIN
if(NEWIND <=1) {
  int POP = 1 + R::rbinom(1,0.2);
  double myETA = R::rnorm(0,sqrt(0.09));
}

double CL = TVCL;

if(POP==2) CL = TVCL * THETA1;

double CLi = CL*exp(myETA);

$CAPTURE POP CL CLi
'

mod <- mcode_cache("C", code)
```


And simulate again
```{r}
set.seed(444)
out <- mrgsim(mod,nid=10000, end=72, events=e,obsonly=TRUE)
```

```{r}
head(out)
mean(out$CL==0.5)
```
```{r}
out %>% 
  filter(time==0) %>%
  group_by(POP) %>% 
  summarise(N=n(), Median = median(CLi))
```


# Three (or more) Populations
There are probably several ways to simulate three populations.  Here is 
one way.  We'll drop the PK model for now and focus on 
generating `POP`.


```{r}
code <- '
$PARAM p1 = 0.33, p2 = 0.6

$PLUGIN Rcpp

$MAIN
if(NEWIND <=1) {
  double mixv = R::runif(0,1);
  int POP = 1;
  if(mixv > p1) POP = 2;
  if(mixv > (p1+p2)) POP = 3;
}

$CAPTURE POP mixv
'
```
Here's what we did
  
  - Code mixture probabilities in `$PARAM`
  - Draw a variate (`mixv`) from `uniform(0,1)`
  - Determine `POP` based on the probabilites and `mixv`
  - Remember: you must use `$PLUGIN` for this to work


Now, let's compile and test it out
```{r}
mod <- mcode_cache("D", code)

set.seed(333)
out <- mrgsim(mod, nid=10000, end=-1)
```

```{r}
head(out)
```
And check that the population is properly configured

```{r}
out %>% as.tbl %>% count(POP) %>% mutate(p = n/nrow(out))
```

And we get back the 33% in population 1, 60% in population 2, and the 
remaining 7% in population 3.

As a final note: remember to call `set.seed()` prior to simulating anything
random with `mrgsovle` in order for the results to be reproducible.

