---
title: "Resimulate random effect variates on demand"
output: github_document
---

# simeta (and simeps)
 
 
```{r,echo=FALSE,message=FALSE}
knitr::opts_chunk$set(comment='.', fig.path = "img/simeta-", message = FALSE)
```

```{r,message=FALSE}
library(mrgsolve)
library(dplyr)
```



## `simeta` example
 
- In this example, we want to simulate a patient-specific baseline response
that is between 80 and 120.
- In the code, we start a loop that calls `simeta` with no 
arguments until the baseline is between the specified bounds
- For this example, we only calculate
`BASE` when `NEWIND <=1` ... or whenever we are working on the 
first record of an individual.  This ensures that we don't re-simulate
`BASE` at every simulation record.
- We have also implemented a counter (`i`) that ensures we only
try to resimulate 100 times.  If a value for `BASE` cannot be
generated in 100 tries, we give up.  
- You probably won't need to implement `FLAG` for your problem. I am 
only using `FLAG` here so we can selectively call the `simeta` code
to demonstrate how it is working.
 


```{r}
code <- '
$PARAM TVBASE = 100, FLAG = 0

$CMT RESPONSE

$MAIN 

if(NEWIND <=1) {

  capture BASE = TVBASE*exp(EBASE);

  int i = 0;

  if(FLAG > 0) {
    while((BASE < 80) || (BASE > 120)) {
      if(++i > 100) {
        report("There was a problem simulating BASE");
      }
      simeta();
      BASE = TVBASE*exp(EBASE);
    }
  }
  
  RESPONSE_0 = BASE;
}


$OMEGA @labels EBASE
1

$CAPTURE EBASE
'
```


```{r,message=FALSE}
mod <- mcode("simeta", code)
```


### Simulate without `simeta`
```{r,comment='.'}
system.time({
  out <- mod %>% mrgsim(nid=100, end=-1)
  sum <- summary(out)
})


print(sum)

```


 
When we simulate with `FLAG=0`, the `simeta` code __isn't__ called and we
`BASE` values all over the place.
 

### Simulate with `simeta`
```{r,comment='.'}
system.time({
  out <- mod %>% mrgsim(nid=100, end=-1, param=list(FLAG=1))
  sum <- summary(out)
})
print(sum)

```

When we simulate with `FLAG=1`, the `simeta` code __is__ called and we
`BASE` values within the specified bounds.
 

## `simeps` example
 
- In this example, we want to re-simulate the residual error variate to 
make sure we have a concentration that is positive.
- We set up a loop that looks like the `simeta` example, but 
we work in `$TABLE` this time because we are calculating `CP`.

 
```{r}
code <- '
$PARAM CL = 1, V = 20, FLAG=0

$SIGMA 50

$PKMODEL cmt="CENT"

$TABLE
capture CP = CENT/V + EPS(1);

int i = 0;
while(CP < 0 && FLAG > 0) {
  if(++i > 100) {
    report("Problem simulating positive CP");
  }
  simeps();
  CP = CENT/V + EPS(1);
}

'
```


```{r}
mod <- mcode("simeps", code)
```


### Simulate without `simeps`

```{r,comment='.'}
system.time({
  out <- mod %>% ev(amt=100) %>% mrgsim(end=48)
  sum <- summary(out)
})

print(sum)

```


__Negative__ concentrations are simulated when we __don't__ call the
`simeps` loop.
 

### Simulate with `simeps`

```{r}
system.time({
  out <- mod %>% ev(amt=100) %>% mrgsim(end=48, param=list(FLAG=1))
  sum <- summary(out)
})

print(sum)
```


 
Better ... all concentrations are positive.
 

