---
title: Sobol sensitivity analysis using sensobol
author: "Kyle Baron"
date: "`r Sys.time()`"
output:
  github_document:
    toc: TRUE
---

```{r, echo = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, 
                      comment = '.', fig.path = "img/sensobol")
```


# Tools
```{r}
library(mrgsolve)
library(tidyverse)
library(PKPDmisc)
library(data.table)
library(sensobol)
```

# The sunitinib PK model
```{r}
mod <- mread("sunit", "model") %>% 
  update(end = 24, delta = 1) %>% zero_re
```

```{r}
see(mod)
```

## Sunitinib dosing 

We are just looking at a single dose for now.

```{r}
sunev <- function(amt = 50,...) ev(amt = amt, ...)
```

# Generate samples

__Number__
```{r}
N <- 2 ^ 15
```



__Generate__

```{r}
mat <- sobol_matrices(N = N, params = c("TVCL", "TVVC", "TVKA", "TVQ", "TVVP"))
mat <- as_tibble(mat)
head(mat)
```


__Transform and groom__

For this example, we will just take uniform samples based on the logarithm of
the current value of the parameter. 
```{r}
params <- unlist(as.list(param(mod))[c("TVCL", "TVVC", "TVKA", "TVQ", "TVVP")])
umin <- log(params / 5)
umax <- log(params * 5)
umin
```

After setting the minimum and maximum for each parameter, get the value of the 
parameter by looking at the quantiles of the (log uniform) distribution


```{r}
mat2 <- imodify(mat, ~ qunif(.x, umin[[.y]], umax[[.y]]))
mat2 <- mutate(mat2, across(everything(), exp), ID = row_number())
head(mat)
```

# Run the analysis

## Simulation

```{r}
out <- mrgsim_ei(mod, sunev(), mat2, output = "df")
out <- as.data.table(out)
```

## Calculate AUC

```{r}
y <- out[, list(auc = auc_partial(time,CP)), by = "ID"][["auc"]]
```

## Indices

```{r}
ind <- sobol_indices(Y = y, N = N, params = names(params), boot = TRUE, R = 1000, first = "jansen")
ind.dummy <- sobol_dummy(Y = y, N = N, params = names(params), boot = TRUE, R = 1000)
```

```{r}
ind
```


## Visualize

```{r}
plot(ind, dummy = ind.dummy) + ylim(0,1)
```
