---
title: Graphical sensitivity analysis 
author: "Kyle Baron"
date: "`r Sys.time()`"
output:
  github_document:
    toc: TRUE
---

```{r, echo = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, 
                      comment = '.', fig.path = "img/graphical-sa")
```

# Introduction 

What I mean by "graphical" is generating simulated outputs with parameters 
varied across a grid of values and making a graphical display of the results.
This is an informal method for visualizing variability in outputs in relation 
to variability in inputs. It is similar to 
[local sensitivity analysis](local-sa.md) in that parameters are varied 
one at a time, but parameters may be varied across the different segments of the 
parameter space. 

This is in contrast to [global sensitivity analysis](global-sensobol.md), 
when all parameters are varied simultaneously across the entire parameter space. 

This vignette uses the mrgsolve in coordination with the 
[mrgsim.sa](https://cran.r-project.org/package=mrgsim.sa) package
package in do local sensitivity analysis.


# Tools
```{r setup}
library(mrgsolve)
library(tidyverse)
library(mrgsim.sa)
```

# Model

We'll perform graphical sensitivity analysis on a 2-compartment PK model.

```{r}
mod <- modlib("pk2") %>%
  update(end = 24, delta = 0.1, outvars = "CP,PERIPH") %>% 
  param(Q = 12, V3 = 50)
```

The parameters of the model are

- CL
- V2
- Q
- V3
- KA

```{r}
see(mod)
```

# Intervention

We are just looking at a single dose for now.

```{r}
dose <- function(amt = 1000,...) ev(amt = amt, ...)
```


# Graphical sensitivity analysis

First, let's look at how to do this with the `mrgsim.sa` package. As a test run
let's vary CL with 30% coefficient of variation around the nominal value 
in the model object

```{r}
ans <- 
  mod %>% 
  ev(dose()) %>% 
  parseq_cv(CL, .cv = 30) %>%
  sens_each() 
```

The output is simulated data in long format

```{r}
head(ans)
```

And we can make a plot of this data
```{r}
ans %>% sens_plot("CP")
```

In the plot
- Parameters are varied across 2 standard deviations of a distribution with 
  30\% CV (see `?parseq_cv`)
- We have some parameters with "low" value and some with "high" value
- There is a dashed reference line that shows the simulated simulated data with 
  no change in the model parameters from what is currently in the model object
  

To see the values in the legend
```{r}
sens_plot(ans, "CP", grid = TRUE)
```


We can look at multiple parameters, this time with larger CV

```{r, fig.width = 8, fig.height = 6}
mod %>% 
  ev(dose()) %>% 
  parseq_cv(CL, V3, KA, Q, .cv = 50) %>%
  sens_each() %>%
  sens_plot("CP")
```


Rather than looking across a certain CV, we can just specify the range
```{r}
mod$CL
mod$V3
mod %>% 
  ev(dose()) %>% 
  parseq_range(CL = c(0.5,2), V3 = c(25,75), .n = 10) %>%
  sens_each() %>%
  sens_plot("CP")
```

Or vary a parameter by a "factor" (double / half)

```{r}
mod %>% 
  ev(dose()) %>% 
  parseq_fct(CL, V3, .n = 6) %>%
  sens_each() %>%
  sens_plot("CP")
```

The user can also simulate across a complete grid of values

```{r, fig.width = 8, fig.height = 7}
mod %>% 
  ev(dose()) %>% 
  parseq_fct(CL, V3, KA, .n = 4) %>%
  sens_grid() %>%
  sens_plot("CP")
```


# Code your own

The `mrgsim.sa` package makes it convenient to simulate and plot these results. 
However, you can also code your own simulations to look at similar types of 
questions. I'll use the chunks below to illustrate the workflow. You'll see that
the output isn't as refined as what you get from `mrgsim.sa`, but it can be 
polished with additional coding. 


To do this sort of sensitivity analysis, we use `idata_set`

```{r}
idata <- tibble(CL = seq_even(0.5, 2, 4))
idata
```

```{r}
out <- mrgsim_ei(mod, dose(), idata, carry_out = "CL", output = "df")
head(out)
```

```{r}
ggplot(out) + geom_line(aes(time, CP, color = factor(CL))) + theme_bw()
```


The code that `mrgsim.sa` uses is just a more  robust or sophisticated 
implementation of the code here. 


