---
title: Local sensitivity analysis using mrgsim.sa
author: "Kyle Baron"
date: "`r Sys.time()`"
output:
  github_document:
    toc: TRUE
---

```{r, echo = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, 
                      comment = '.', fig.path = "img/local-sa")
```

# Introduction 

Local sensitivity analysis consists of comparing variation in model parameters
against variation in model out puts when the parameters are varied one at a
time and by only very small amounts. This is in contrast to 
[global sensitivity analysis](global-sensobol.md), when all parameters are 
varied simultaneously across the entire parameter space. 

This vignette uses the mrgsolve in coordination with the 
[mrgsim.sa](https://cran.r-project.org/package=mrgsim.sa) package
package in do local sensitivity analysis.


# Tools
```{r setup}
library(mrgsolve)
library(tidyverse)
library(mrgsim.sa)
```

# Model

We'll perform local sensitivity analysis on a 2-compartment PK model.

```{r}
mod <- modlib("pk2") %>%
  update(end = 72, delta = 0.1, outvars = "CP") %>% 
  param(Q = 12, V3 = 50)
```

The parameters of the model are

- CL
- V2
- Q
- V3
- KA

```{r}
see(mod)
```

# Intervention

We are just looking at a single dose for now.

```{r}
dose <- function(amt = 1000,...) ev(amt = amt, ...)
```


# Local sensitivity analysis

The function we'll use is `mrgsim.sa::lsa`. The  call to `lsa` looks like

```{r}
res <- lsa(mod, par = "CL,V2,Q,V3,KA", var = "CP", events = dose())
```

We have specified 

- `par`: the names of the parameters we want to include in the sensitivity 
  analysis
- `var`: the name of the output variable; the `mrgsolve` model must generate 
  this output
- `events`: this is passed under the `...` argument to `lsa()`; these arguments 
  eventually get passed to the simulation function (see below)

We have not specified `fun`, which is a non-exported wrapper function for 
`mrgsolve::mrgsim()`

```{r}
mrgsim.sa:::.lsa_fun
```

You can create your own `fun`ction and then pass objects to that function 
to carry out the simulation. Note that `lsa()` will always pass `mod`
to your simulation function as the first argument.

## Visualize

The local sensitivity analysis usually completes very quickly. The output 
is a regular data frame with the simulated sensitivity data in long 
format

```{r}
head(res)
```

`mrgsim.sa` provides a plot method for visualizing the result

```{r}
res %>% filter(time <= 72) %>% plot()
```

Or you can easily make your own plot

```{r}
ggplot(res, aes(time, sens, color = p_name)) + 
  facet_wrap(~p_name) + 
  geom_line(lwd=1) + 
  theme_bw() + theme(legend.position = "top")
```


## Multiple outputs

For example, we can look at multiple outputs from an indirect response model

```{r}
mod <- modlib("irm1", delta = 0.1, end = 14, rtol = 1e-10) 
```

```{r}
param(mod)

mrgsim(mod, dose()) %>% plot("CP,RESP")
```

Run the analysis 

```{r}
res <- lsa(mod, par = "CL,V2,Q,KA,IC50", var = "CP,RESP", events = dose())
```

Plot the result

```{r}
plot(res)
```
