---
title: Simulating large time-to-event trials with mrgsolve
output: github_document
---

# Introduction

This vignette shows implementation of MTIME approach to
simulating large time-to-event trial in NONMEM. We show the 
Repeated Time-to-Event model. 

## Reference

__Simulating large time-to-event trials in NONMEM__

Joakim Nyberg, Kristin E. Karlsson, Siv Jönsson, Ulrika S.H. Simonsson, Mats O. Karlsson and Andrew C. Hooker


PAGE 23 (2014) Abstr 3166 
https://www.page-meeting.org/?abstract=3166


```{r, message = FALSE, warning = FALSE}
library(tidyverse)
theme_set(theme_bw())
library(mrgsolve)
```

# Input data

```{r}
#| comment: '.'
data <- data.frame(
  TIME = c(0, 288), 
  CMT = 0, 
  AMT = 0, 
  EVID = 0
)
ids <- seq(1, 300)
data <- crossing(ID = ids, data)

head(data)
```

# RTTE Model
```{r}
#| comment: '.'
mod <- mread("rtte.txt", end = -1)

set.seed(98765)
out <- mrgsim(mod, data)

head(out)
```

# Check event output
```{r}
#| comment: '.'

res <- as.list(mod@envir)
x <- matrix(res$data, ncol = length(res$names), byrow=TRUE)
x <- as.data.frame(x)
names(x) <- res$names

head(x)
```


## Calculate analytic Event Time

```{r}
x <- mutate(
  x, 
  AT = -log(R)/BASE, 
  DIFF = TIME - AT
)

head(x)
```

## Difference between MTIME and Analytic Time

Take time to first event

```{r}
xx <- distinct(x, ID, .keep_all = TRUE)
ggplot(xx, aes(x = DIFF)) + 
  geom_histogram(color = "white")
```

# Model code

```{r, eval = FALSE, code = mod@code}

```

