---
title: Fit a model with multiple endpoints
output:
  github_document:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(comment = '.', message = FALSE, warning = FALSE)
```

# Setup 
```{r}
library(mrgsolve)
library(tidyverse)
theme_set(theme_bw())
```


# Exploratory data analysis

## Concentration versus time

We have plasma concentrations fairly early after the dose and 
runing out to 48 hours:

```{r}
conc <- readRDS("two_endpoints_plasma.RDS")

conc

ggplot(conc, aes(time,CP)) + 
  geom_point() + geom_line() + 
  ylab("Plasma concentration") + ylim(0,5)
```

## Cumulative amount in urine versus time

```{r}
urine <- readRDS("two_endpoints_urine.RDS")

urine

ggplot(urine, aes(time,UR)) + 
  geom_point() + geom_line() + 
  ylab("Cumulative amount in urine") + 
  xlim(0,48)
```

# The model

The data above were simulated with this model, including the parameter
values you see below.

```{r}
mod <- mread("two_endpoints.cpp")
```

```{c, eval=FALSE, code = mod@code}

```

Check out the `[ TABLE ]` block.  We check the value of `dvtype` and 
return concentration if `dvtype==1` and urine amount if `dvtype==2`.  

Note that there isn't really a loss of generality here.  You can 
still simulate from the model as you wish.  In order for `DV` to be 
meaningful, you'll have to use a data set with `dvtype` defined.  But
you could still use this model to simulate any compartment / output
at any time ... just ignore `DV` and access `CP` or `UR`.


# Assemble a model estimation data set

## Combine concentration and urine data
First, add a `DV` column to both data sets and set a flag
for the type of each `DV`
```{r}
conc <-  mutate(conc,  DV = CP, dvtype = 1)
urine <- mutate(urine, DV = UR, dvtype = 2)
```


Then make a single data set sorted by time of observation
```{r}
obs <- bind_rows(conc,urine) %>% arrange(time)
```

No longer need `CP` or `UR`
```{r}
obs <- select(obs,ID,time,dvtype,DV)
```

And we'll add some columns that to make mrgsolve happy; this is the
observations part of the data set
```{r}
obs <- mutate(obs, evid = 0, cmt = 0, amt = 0)

obs
```

## Add a dosing record
Just like we did in the `Indometh` example, pull a single row
and set up the dosing

```{r}
dose <- obs %>% slice(1)

dose
```

Now, set up this chunk to be doses (not observations)

```{r}
dose <- mutate(dose, time = 0, amt = 100, evid = 1, cmt = 1)
```

We make this `dvtype` 0 and set `DV` to `NA` (it is dosing record
so we don't want this to contribute information to parameter 
estimates)

```{r}
dose <- mutate(dose, dvtype = 0, DV = NA)
```


Then bind it all together and arrange

```{r}
data <- bind_rows(obs,dose) %>% arrange(time)

data
```

# Estimate parameters: Approach I

## Objective function

```{r}

ofv1 <- function(p, data, yobs, dvcol = "DV", pred=FALSE) {
  p <- lapply(p, exp)
  names(p) <- names(theta)
  mod <- param(mod,p)
  out <- mrgsim_d(mod, data, output="df")
  if(pred) return(as_tibble(out))
  y_hat <- out[[dvcol]]
  sum((y_hat-yobs)^2,na.rm=TRUE)
}
```

## Optimize
```{r}
yobs <- data[["DV"]]

theta <- log(c(CLnr = 2, V=10, KA = 1, CLr = 1))

fit <- minqa::newuoa(theta, ofv1, data=data, yobs=yobs)
```


## Results

The way we set this up, the true parameters are in the model

```{r}
as.numeric(param(mod))
```

```{r}
exp(fit$par)
```


And make the plots for concentration and urine cumulative amount; 
we use the `dose` data set (above) to get the smooth line

```{r}
prd <- ofv1(fit$par,data = dose, pred = TRUE)
```


```{r}
ggplot() + 
  geom_point(data=conc, aes(time,CP), size = 4) + 
  geom_line(data=prd, aes(time,CP),col="firebrick", lwd=1) +
  scale_y_log10() + ylab("Plasma concentration")

ggplot() + 
  geom_point(data= urine, aes(time,UR), size = 4) + 
  geom_line(data=prd, aes(time,UR),col="cornflowerblue", lwd=1) +
   ylab("Cumulative amout in urine")

```


# Estimate parameters: Approach II

I still think approach I is simpler and easier to pull off without 
loss of flexibility for simulation.

But if you didn't want to do the `DV` approach, here is another way.


## Some more data grooming

First, find which rows are type 1 (concentration) and which 
are type 2 (urine):

```{r}
iconc <- which(data$dvtype==1)
iur   <- which(data$dvtype==2)
```

So we can use this `iur` vector to grab just the urine amounts
```{r}
iur

data[iur,]
```

Now, create a vector of all the concentrations followed by all 
the urine amounts; the final product is `yobs`

```{r}
yplas <- data$DV[iconc]
yur   <- data$DV[iur]
yobs  <- c(yplas,yur)
```

## The objective function

The OFV function will now have to modified as well

```{r}
ofv2 <- function(p,data,pred=FALSE) {
  p <- lapply(p, exp)
  names(p) <- names(theta)
  mod <- param(mod,p)
  out <- mrgsim_d(mod,data)
  if(pred) return(as_tibble(out))
  
  ## The action is here:
  yplas_hat <- out$CP[iconc]
  yur_hat   <- out$UR[iur]
  y_hat     <- c(yplas_hat,yur_hat)
  ##---------------------------
  
  sum((y_hat-yobs)^2)
}
```

The function is mostly the same, but now we'll slice the simulation 
output to first grab the concentrations and then grab the urine
concentrations.  __NOTE__: this works because the simulation output
has the exact same design / setup as the input data.  So we slice both 
the input data and the output by the same indices so we can be sure 
the observed and simulated values match up.

## Optimize

Ok, now try it out:

```{r}
theta <- log(c(CLnr = 2, V =12, KA = 1, CLr = 1))

fit <- minqa::newuoa(par=theta, fn=ofv2, data=data)
```

## The results

```{r}
as.numeric(param(mod))

fit$par
```


And plot
```{r}
prd <- ofv2(fit$par,data = dose, pred = TRUE)

ggplot() + 
  geom_point(data= data[iconc,], aes(time,DV), size = 4) + 
  geom_line(data=prd, aes(time,CP),col="firebrick", lwd=1) +
  scale_y_log10() + ylab("Plasma concentration")

ggplot() + 
  geom_point(data= data[iur,], aes(time,DV), size = 4) + 
  geom_line(data=prd, aes(time,UR),col="cornflowerblue", lwd=1) +
  scale_y_log10() + ylab("Cumulative amout in urine")

```

```{r,include = FALSE}
saveRDS(file = "two_endpoints_data.RDS", data)
```

