---
title: "Generating input data sets for mrgsolve"
output: github_document
---


```{r,echo=FALSE,message=FALSE}
knitr::opts_chunk$set(comment='.',fig.path="img/input_data_sets-",message=FALSE)
```

```{r,message=FALSE}
library(mrgsolve)
library(dplyr)
```



## Input data sets
An important mechanism for creating robust, complex simulations is
the input data set.  Input data sets specify the population 
of individuals to simulate, including the number of individuals,
each individual's dosing interventions, each individual's covariate
values etc.  The input data set is just a plain old `R` `data.frame`, but 
with some expectations about which columns are present and expectations 
for how to handle columns for certain names.  For example, 
every input data set has to have an `ID`, `time`, and `cmt` column. Note
that either lower case names (like `time` and `cmt`) are acceptable
as are upper case names (like `TIME` and `CMT`).  But users are not
to mix upper and lower case names (like `time` and `CMT`) for certain
column names related to dosing events.  The help topic `?data_set` 
discusses more about what the expectations are for input data sets.

<hr>
## Functions to generate input data sets
`mrgsolve` provides several functions and workflows to help you 
put together the right input data set for your simulation. The main point 
of this blog post is to review some of these functions to help you 
better organize your `mrgsolve` simulations.  Some functions are very 
simple and you might not find a function to do __exactly__ what you 
want to do.  But we've found these functions to be helpful to 
accomplish tasks that we found ourselves repeating over and over ... and 
thus these tasks were formalized in a function.  Just keep in mind 
that input data sets are just `data.frames` ... you can use any code 
or any function (even your own!) to do tasks similar to what these functions are doing.
<hr>
### `expand.ev`

`expand.ev` is like `expand.grid`: it creates a single `data.frame` with all
combinations of it's vector arguments.  It's pretty simple but convenient to 
have.  For example,

```{r}
data <- expand.ev(amt=c(100,200,300), ID=1:3)
data
```

This function call gives us 3 individuals at each of 3 doses.  The `expand.grid` nature of `expand.ev` is what gives us `3x3=9` rows in the 
data set.  Notice that the `IDs` are now 1 through 9 ... `expand.ev` renumbers `IDs` so that there is only one dosing event per row and 
there is on row per `ID`. 

Also notice that `time` defaults to 0, `evid` defaults to 1, and `cmt` defaults to 1.  So, `expand.ev` fills in some of the required columns
for you. 

Let's simulate with this data set:
```{r}
mod <- mrgsolve:::house() %>% Req(CP) 
mod %>% 
  mrgsim(data=data) %>% 
  plot(CP~time|factor(ID),scales="same")
```

<hr>
### `as_data_set`
This function allows you to combine several event objects into a single data sets.  An example works best to illustrate.


First, create three event objects.  Let's try one `ID` at 100 mg, 
two `IDs` at 200 mg, and 3 `IDs` at 300 mg.
```{r}
e1 <- ev(amt=100, ID=1)
e2 <- ev(amt=200, ID=1:2)
e3 <- ev(amt=300, ID=1:3)
```

The events are
```{r}
e1
```
and
```{r}
e2
```
and
```{r}
e3
```

When we combine these events with `as_data_set` we get
```{r}
data <- as_data_set(e1,e2,e3)
data
```

A nice feature of `as_data_set` is, unlike `expand.ev` and the previous example, we can use complicated event sequences that are expressed with
more than one line in the data set.  For example, consider the case
where every `ID` gets a 250 mg loading dose, and then either get 250 mg q24h, or 120 mg q12h or 500 mg q48h.

```{r}
load <- function(n) ev(amt=250, ID=1:n)
e1 <- load(1) + ev(amt=250, time=24, ii=24, addl=3, ID=1)
e2 <- load(2) + ev(amt=125, time=24, ii=12, addl=7, ID=1:2)
e3 <- load(3) + ev(amt=500, time=24, ii=48, addl=1, ID=1:3)
```

Now, `e1`, `e2`, and `e3` are more complex
```{r}
e1
e3
```

But, we can still pull them together in one single data set
```{r}
data <- as_data_set(e1,e2,e3)
data
```

An example simulation
```{r}
set.seed(1112)
mod %>% 
  omat(dmat(1,1,1,1)/10) %>% 
  data_set(data) %>% 
  mrgsim() %>% 
  plot
```
<hr>
### `as.data.frame.ev`
Just a quick reminder here that you can easily convert between 
a single event object and a `data.frame`
```{r}
as.data.frame(e3)
as.ev(as.data.frame(e3))
```

So if you were building up an event object and just wanted to 
use it as a `data_set` or as a building block for a `data_set`, just
coerce with `as.data.frame`.

<hr>
### `assign_ev`
This function assigns an intervention in the form of an event object
to individuals in an `idata_set` according to a grouping column.

To illustrate, make a simple `idata_set`
```{r}
set.seed(8)
idata <- data_frame(ID=sample(1:6), arm=c(1,2,2,3,3,3))
idata
```

Here, we have 6 `IDs`, one in arm 1, two in arm 2, three in arm 3.  Let's take the events from the previous example and assign them to the different 
arms.

```{r}
e1 <- ev(amt=250) + ev(amt=250, time=24, ii=24, addl=3)
e2 <- ev(amt=250) + ev(amt=125, time=24, ii=12, addl=7)
e3 <- ev(amt=250) + ev(amt=500, time=24, ii=48, addl=1)
assign_ev(list(e3,e2,e1),idata,"arm")
```

Please look carefully at the input (`idata` and `list(e3,e2,e1)`); I have mixed
it up a bit here to try to illustrate how things are assigned.  



<hr>
### `ev_days`
This is a recently-added function (hint: you might need to install the latest
version from GitHub to use this) that lets you schedule certain events on certain
days of the week, repeating in a weekly cycle.  

For example, to schedule 250 mg doses every Monday, Wednesday, and Friday for a
month, you can do
```{r}
data <- ev_days(ev(amt=250, ID=1), days="m,w,f", addl=3)
data

mod %>% mrgsim(data=data,end=168*4) %>% plot
```


Or, you can do 100 mg doses on Monday, Wednesday, Friday, and 50 mg doses on Tuesday, Thursday, with drug holiday on weekends
```{r}
e1 <- ev(amt=100,ID=1)
e2 <- ev(amt=50,ID=1)
data <- ev_days(m=e1,w=e1,f=e1,t=e2,th=e2,addl=3)
data
```

And simulate
```{r}
mod %>% mrgsim(data=data,end=168*4) %>% plot
```


The same thing can be accomplished like this
```{r}
a <- ev_days(e1,days="m,w,f",addl=3)
b <- ev_days(e2,days="t,th",addl=3)

c(as.ev(a),as.ev(b))
```
You can use this as an event object or just coerce to `data.frame` to use as
a `data_set`.
<hr>
## Filter input data set inline
Remember, when you pass in your input data set via `data_set`, you can 
filter in line:

```{r}
data <- expand.ev(amt=c(100,200,300))
mod %>% data_set(data, amt==300) %>% Req(GUT,CP) %>% mrgsim
```








