---
title: "Parallel zero-/first-order absorption"
output: 
  github_document
---

```{r,echo = FALSE, message = FALSE}
knitr::opts_chunk$set(fig.path = "img/parallel-", comment='.', 
                      message = FALSE, warning = FALSE)
```



```{r}
library(dplyr)
library(mrgsolve)
```

## PK model

- One-compartment (`CENT`), with absorption / depot compartment (`GUT`)
- We want to administer a single dose, with a certain fraction as a
  a zero-order infusion and the remaining fraction by first order absorption
- To split to dose, we use the bioavailability parameter for `GUT` and `CENT` 
  (`F_GUT` and `F_CENT`, respectively) and we specify the duration of the infusion
  with `D_CENT`.


```{r}
code <- '
$PARAM CL = 1, V = 20, KA = 1.1
frac_infusion = 0.37, dur_infusion = 10

$PKMODEL cmt = "GUT CENT", depot = TRUE

$MAIN

D_CENT = dur_infusion;
F_CENT = frac_infusion;
F_GUT =  1-frac_infusion;

$TABLE
capture CP = CENT/V;
'
```



```{r}
mod <- mcode("parallel", code) %>% update(end = 48, delta = 0.1)
```

## Parallel zero-/first-order

- We have 100 mg total dose
- We will need an event object (or data set) with two dosing records, giving 
  the full dose on each record every time we want to administer a dose

```{r}
infusion <- ev(amt = 100, cmt = "CENT", rate = -2)

zero <- mutate(infusion, cmt = "GUT", rate= 0)

dose <- c(infusion, zero)

dose
```
# Result


```{r}
out <- mrgsim_e(mod, dose)
```


```{r}
plot(out, GUT + CP ~time)
```


## Show individual components

Let's set up a simulation where we have 3 subjects: one gets the infusion 
(only; blue line), one gets the depot dose (only; pink line) and the other gets
both (green line):

```{r}
data <- as_data_set(infusion, zero, dose)

data
```


```{r}
mrgsim_d(mod,data) %>% plot(GUT+CP~time)
```







