---
title: "Time-varying KA"
output:  github_document
---

```{r,echo = FALSE, message = FALSE}
knitr::opts_chunk$set(fig.path = "img/time-varying-", comment='.', 
                      message = FALSE, warning = FALSE)
```


This gallery piece shows now to implement a time-varying, first-order
absorption rate constant.


```{r}
library(dplyr)
library(mrgsolve)
```



Absorption will be first order, and start out very fast (large KA) and then 
slow down after a while (smaller KA).

The main things we need to do:   

1. Reset everything when it's a new individual
1. When there is a dose, switch to the fast KA and set the time when KA will 
   revert back to slow
1. At the change time, revert to slow

```{r}
code <- '
[ param ] KAfast = 3, KAslow = 0.5, CL = 1, V = 5

[ pkmodel ] cmt = "GUT CENT", depot = TRUE

[ main ]

if(NEWIND <=1) {
  double KA = KAslow;
  double mt = -1;
}

if(EVID==1) { 
  KA = KAfast;
  mt = self.mtime(TIME + 0.2);
}

if(TIME > mt) { 
  KA = KAslow;
}

[ table ] capture CP = CENT/V;
'
```

Set up a simulation from this model where we look at all slow absorption (blue 
line) or fast then slow (pink line)
```{r}
mod <- mcode("foo", code) %>% update(outvars = "CP")

idata <- tibble(KAfast = c(mod$KAslow,2.5))
```

```{r}
out <- mrgsim_ei(mod, ev(amt = 100), idata = idata, end = 12, delta = 0.2)

plot(out,logy=TRUE)

```










