---
title: "Simple adaptive dosing"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment='.', message=FALSE, warning=FALSE)
```


# Introduction

This is a very simple dose adjustment example, presented as a first example 
to get started.  

The scenario is:

1. Dosing regimen of 1200 mg BID over 10 days
1. Sensitivity analysis for volume of distribution (10 to 40 L)
1. If the concentration ever reaches some critical concentration (say 100 mg/L), 
   we cut the dose by some factor (say in half)
1. Once the dose is adjusted it remains adjusted
1. Evaluate time when dose adjustment takes place

# The model

```{r}
library(mrgsolve)
library(dplyr)

mod <- mread("adaptive_simple", end = 240)
```


```{c, eval=FALSE, code = mod@code}

```

## To note

1. We set `condition_met` to `false` in preamble and then renew that in 
   `[ main ]` when a new individual is encountered.
1. The bioavailability fraction is adjusted by some amount if the critical 
   condition is met.
1. We always make the adjustment in `[ table ]`, that is, after the system has
   advanced so that doses are adjusted starting the __next__ time the system 
   advances

# Simulate

```{r}
data <- expand.ev(amt = 1200, ii = 12, addl = 19, V = c(10, 20, 30, 40))

out <- mrgsim(mod, data, carry_out = "V")

plot(out, CP ~ time | factor(V), scales="same", layout=c(4,1))

```


```{r}
plot(out, F_DEPOT ~ time, scales = "same")
```


# Important

This very simple model cuts the dose in half when a critical concentration is 
seen.  Whether or not the critical concentration is seen depends on the sampling
scheme (when observations are made).  For most applications, I have found it 
necessary to explicitly state __when__ the critical dose adjustment observation 
is made (e.g. at the trough or at week 14 or whatever).  This most likely 
requires some special handling in the data set, usually including records
in the data set with a "custom" `evid` (like 33) that is your signal for 
checking the state of the model and adjusting the dose. mrgsolve won't do 
anything with `evid` 33, but you can

For example:

```{c, eval=FALSE}
[ table ]
condition_met = ((EVID==33) && (CP > condition)) || condition_met;
```

