---
title: "Sigma-Like THETA"
output: 
  github_document
---

```{r,echo = FALSE, message = FALSE}
knitr::opts_chunk$set(fig.path = "img/sigma-like-theta-", comment='.', 
                      message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(mrgsolve)
```

In this model, we will model the residual error variances with parameters 
(or THETAs) rather than SIGMAs. We will also model RUV in the standard way 
(using SIGMAs) so we can compare.

## In the model

- parameter `THETA10` is proportional error
- parameter `THETA11` is additive error
- `SIGMA` 1 is equivalent to `THETA10` (proportional error)
- `SIGMA` 2 is equivalent to `THETA11` (additive error)
- `SIGMA` 3 would be fixed to 1 in the NONMEM run (here, just set to 1)


```{r}
code <- ' 
$PARAM THETA10 = 0.025,  THETA11 = 5, CL = 1, V = 25

$PKMODEL cmt = "CENT"

$SIGMA 0.025 5

$SIGMA 1 // "FIXED"

$TABLE
capture IPRED = CENT/V;

double W = sqrt(THETA11 + THETA10*pow(IPRED,2.0));

capture DV = IPRED + W*EPS(3);

capture DV2 = IPRED*(1+EPS(1)) + EPS(2);
'

```

__Compile__
```{r}
mod <- mcode("sigma-like-theta", code, delta = 0.5, end = 10000) 
```

__Simulate__
```{r}
out <- mrgsim_df(mod, events = ev(amt = 50, ii = 24, addl = 10000))
```

__Check__
```{r}
head(out)
```

__Plot__
```{r}
library(ggplot2)
p <- ggplot(out, aes(sample = DV, distribution = DV2)) +  theme_bw()
p + stat_qq_line(col = "firebrick", lwd = 1) + stat_qq() 
```

__Summarise__

```{r}
out %>% 
  pivot_longer(cols = DV:DV2) %>% 
  group_by(name) %>% 
  summarise(tibble(mean = mean(value), SD = sd(value)), n = n())
```


